### 2025-02-03 data/data_cleaning.Rmd
Initial Upload

Portugal fertility survey 1979

- [More information](https://wfs.dhsprogram.com/index.cfm?ccode=pt)
- [data source](https://wfs.dhsprogram.com/)
- [data dictionary](https://wfs.dhsprogram.com/pt/ptsr01.dct)

File `portugal.RData` on the course web site.  Code in `Assignment1.Rmd`

```{r dataDownload, include=FALSE}
pUrl = 'http://wfs.dhsprogram.com/pt/ptsr01.dat'
pName = file.path(tempdir(), 'portugal.dat')
if(!file.exists(pName)) {
  download.file(pUrl, pName)
}

datNames = rbind(
		age=c(45,2),
		ageMarried=c(149,2), 
		monthsSinceM = c(157,4),
#		failedPregnancies=c(421,2),
#		failedPregStill=c(423,2),
#		failedPregSpAb=c(425,2),
		pregnancies=c(433,2),
		children=c(435,2),
		sons=c(443,2),
#		firstBirthInterval = c(479,2),
		region = c(641,2),
		literacy = c(649,2)
)
		colnames(datNames ) = c('start','len')
		datNames = cbind(startm1=datNames[,1]-1,datNames, sum=apply(datNames, 1,sum))
		cbind(datNames[-1,1] , datNames[seq(1, nrow(datNames)-1),4])
		datNames[-1,1] = datNames[-1,2] - datNames[seq(1, nrow(datNames)-1),4]
		dWidths = as.vector(t(datNames[,c(1,3)]))
		dNames = paste(rep(rownames(datNames), rep(2, nrow(datNames))),
  	rep(c( "junk",""), nrow(datNames)), sep="")

		dNames = dNames[dWidths > 0]
		dWidths = dWidths[dWidths > 0]

		formats = list(
			ageMarried = data.frame(code=1:7,  label=c(0,15,18,20,22,25,30)),
			region = data.frame(code=1:5, 
				label=c('lisbon','porto','20k+', '10-20k', 'lt10k')),
			literacy = data.frame(code=1:2, label=c('yes','no')),
			firstBirthInterval = data.frame(
					code = 1:8,
					label = c(
							'lt0','0-7', '8-11','12-23',
							'24-35','36-47','48-59','60-Inf'
							)
					)
		)

		formats$ageMarried$label = 
  	paste(formats$ageMarried$label, 'to',
  	c(formats$ageMarried$label[-1], 'Inf'), sep='')
  	formats$ageMarried = rbind(formats$ageMarried, data.frame(code=88, label='never'))

  portugal = read.fwf(
    pName,
    dWidths, col.names=dNames,
    header=FALSE)

  portugal = portugal[,grep("junk$", names(portugal), invert=TRUE)]

for(D in intersect(names(portugal), names(formats))){
  		portugal[[D]] = factor(portugal[[D]],
  			levels=formats[[D]]$code, 
				labels=formats[[D]]$label)
}
portugal$ageMarried = relevel(portugal$ageMarried, '22to25')
portugal$region = relevel(portugal$region, 'lt10k')

if(TRUE) save(portugal, file='raw_data.RData')
```

```{r thedata}
head(portugal)
table(portugal$region)
```

Region is `lt10k` rural areas (less than 10,000 people), towns of size 10-20k, 20k+, and the two largest cities (Lisbon and Porto).

The Question:

- How do literacy and age of marriage affect family size?
- After we account for known explanatory variables, how much variation (if any) is there in birth rates?

Notes:

- By European standards Portugal is a poor country, and in 1980 it had the same GDP per capita as Mexico.
- It's well known that families are larger in rural areas (i.e. `region` is a confounder).
- You should proceed as if an expert in this area has told you that you do not need to consider zero-inflation.  Zero-inflated models do not fit well to this data, probably because birth rates are lower than Fiji, many zeros are expected, and the likelihood is flat.