### 2025-02-04 paper/paper.Rmd
Changed File Type

---
title: "Analyzing Influencing Factors of House Prices in Ames by Linear Regression"
author: "Jiadong Wang, Yanzun Jiang"
date: today
date-format: long
thanks: "Code and data supporting this paper is available at: <https://github.com/Stary54264/factors_affect_family_size_in_portugal>"
output: pdf_document
number-sections: true
toc: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.