### 2025-02-10 sta303-hw1.Rmd
update

---
title: "sta303-hw1"
output: html_document
date: "2025-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(MASS)  
library(ggplot2)
library(here)
load(here::here("data/Portugal.RData"))
install.packages("kableExtra", repos = "https://cloud.r-project.org/")

```

```{r}
portugal

summary(portugal)

table(portugal$literacy)
table(portugal$ageMarried)

library(dplyr)
library(knitr)
library(kableExtra)

numerical_summary <- portugal %>%
  select(where(is.numeric)) %>%
  summarise(
    across(
      everything(),
      list(
        Mean = ~ mean(.x, na.rm = TRUE),
        SD = ~ sd(.x, na.rm = TRUE),
        Min = ~ min(.x, na.rm = TRUE),
        Max = ~ max(.x, na.rm = TRUE)
      )
    )
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("Variable", ".value"),
    names_sep = "_"
  )

categorical_summary <- portugal %>%
  select(where(is.factor)) %>%
  summarise(
    across(
      everything(),
      ~ paste(names(table(.x)), "(", table(.x), ")", collapse = "\n")
    )
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Summary"
  )

numerical_summary %>%
  kbl(caption = "Numerical Summary Table for Dataset 'portugal'", align = "l") %>%
  kableExtra::kable_styling(full_width = FALSE, position = "center") %>%
  kableExtra::column_spec(3, width = "20em")

categorical_summary %>%
  kbl(caption = "Categorical Summary Table for Dataset 'portugal'", align = "l") %>%
  kableExtra::kable_styling(full_width = FALSE, position = "center") %>%
  kableExtra::column_spec(2, width = "30em")

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
# Bar plot of ageMarried categories
library(ggplot2)
library(tidyverse)
ggplot(portugal, aes(x = ageMarried)) +
  geom_bar(fill = "red") +
  theme_minimal() +
  labs(title = "Distribution of Age Married", x = "Age Married", y = "Count")

# Boxplot of children by literacy(categorical variable)
ggplot(portugal, aes(x = literacy, y = children)) +
  geom_boxplot(fill = "blue") +
  theme_minimal() +
  labs(title = "Number of Children by Literacy", x = "Literacy", y = "Children")

ggplot(portugal, aes(x = ageMarried, y = children, fill = literacy)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  facet_wrap(~region) +
  theme_minimal() +
  labs(title = "Average Number of Children by Age Married and Literacy",
       x = "Age Married",
       y = "Average Number of Children",
       fill = "Literacy")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
portugal$logYearsMarried = log(pmax(1, portugal$monthsSinceM)/12)
#set reference-"18to20"
portugal$ageMarried = relevel(portugal$ageMarried, "18to20")

poi_model = glm(children ~offset(logYearsMarried) + literacy + ageMarried, data = portugal, family = poisson)

knitr::kable(summary(poi_model)$coef, digits = 3)

library(dplyr)
library(knitr)

model_summary <- summary(poi_model)

coef_summary <- as.data.frame(model_summary$coefficients)

colnames(coef_summary) <- c("Estimate", "Std. Error", "z value", "Pr(>|z|)")

knitr::kable(coef_summary, caption = "Comparison of the log-rate ratios and 95% confidence intervals for Negative Binomial model relating literacy and age at marriage to the number of children, with an offset for years married. The results indicate similar trends as the Poisson model, with literacy being significant, and age at marriage having mixed effects. The reference category for age at marriage is 18 to 20 years.", digits = 3, align = "c") %>%
  kableExtra::kable_styling(full_width = FALSE, position = "center") %>%
  kableExtra::column_spec(1, bold = TRUE)

```

```{r}
poi_modelCI = as.data.frame(exp(confint(poi_model)[-1,]))
poi_modelCI$level = gsub(paste(names(poi_model$xlevels), collapse='|'),
   "", rownames(poi_modelCI))
poi_modelCI$variable=unlist(strsplit(rownames(poi_modelCI),poi_modelCI$level))
poi_modelCI$x = 1:nrow(poi_modelCI)
poi_modelCI$cex = sqrt(1/apply(log(poi_modelCI[,1:2]),1,diff))
forXaxis = tapply(poi_modelCI$x, poi_modelCI$variable, mean)

poi_modelCI[1:3,]
```
```{r}
matplot(poi_modelCI[,1:2], type='n',
   xaxt='n', bty='n', xaxt='n',
   log='y', ylab='RR')

segments(poi_modelCI$x, poi_modelCI[,1],
   poi_modelCI$x, poi_modelCI[,2])

numeric_coef <- as.numeric(as.character(poi_modelCI$coef[-1]))
numeric_coef <- rep(0, length(poi_modelCI$x))  # Ensure the same length

 points(poi_modelCI$x,
   exp(numeric_coef), pch=15,
   cex = poi_modelCI$cex, col = '#00000030')

mtext(poi_modelCI$level, 1, at=poi_modelCI$x,
   las=3, line=-1)

mtext(names(forXaxis), 1, at=forXaxis,
   line=-2)
abline(h=1, lty=3)

mtext(poi_modelCI$level, side = 1, at = poi_modelCI$x, las = 1, line = 3, cex = 0.5)
```

```{r}
#install.packages("TMB", type = "source")
#packageVersion("TMB")
#packageVersion("Matrix")
#.rs.restartR()

library('glmmTMB')
NB_model = glmmTMB(children ~ offset(logYearsMarried) + literacy +
   ageMarried, data = portugal,family=nbinom2)

NB_model_summary <- summary(NB_model)

nb_coef_summary <- as.data.frame(NB_model_summary$coefficients$cond)

colnames(nb_coef_summary) <- c("Estimate", "Std. Error", "z value", "Pr(>|z|)")

knitr::kable(nb_coef_summary, caption = "Summary Table for Model: NB_model", digits = 3, align = "c") %>%
  kableExtra::kable_styling(full_width = FALSE, position = "center") %>%
  kableExtra::column_spec(1, bold = TRUE)

```

```{r}
knitr::kable(
   rbind(confint(NB_model)[1:7,c(3,1,2)],
 sd=1/sqrt(confint(NB_model, parm='sigma'))[c(3,2,1)]), digits=3)
```

```{r}
NB_modelFitCI = as.data.frame(exp(confint(NB_model)[-1,]))
NB_modelFitCI$level = gsub(paste(names(NB_model$xlevels), collapse='|'),
   "", rownames(NB_modelFitCI))
NB_modelFitCI$variable=unlist(strsplit(rownames(NB_modelFitCI),NB_modelFitCI$level))
NB_modelFitCI$x = 1:nrow(NB_modelFitCI)
NB_modelFitCI$cex = sqrt(1/apply(log(NB_modelFitCI[,1:2]),1,diff))
forXaxis = tapply(NB_modelFitCI$x, NB_modelFitCI$variable, mean)

NB_modelFitCI[1:3,]
```

```{r}

```

```{r}
exp(confint(NB_model, 1))
1/sqrt(confint(NB_model, parm = "sigma"))
```